Topics in Algorithms

### 2018-01-10 project/paper.tex

Furthermore $d_{G}(p_{i_1}(u),u) \leq i_1 \cdot d_G(u,v)$. When $p_j(u) = w \not\in B_v$
we have $d_G(v,p_{j+1}(v)) \leq d_G(v,p_{j}(u)) \leq d_G(v,u) + d_G(u,p_j(u))$ by Lemma \ref{lem:onedeltaperiteration} and the triangle inequality. Hence each iteration of $dist_k(u,v,i)$ increases $d_G(u,w)$ by at most $d_G(u,v)$. At termination we have
\begin{align*}
d_G(u,w) + d_G(w,v) &\leq 2d_G(u,w) + d_G(u,v) \\
&\leq 2(i_1 \cdot d_G(u,v) + (i_2 - i_1)d_G(u,v)) + d_G(u,v) \\
&\leq 2 \cdot i_2 \cdot d_G(u,v) + d_G(u,v) \\
&\leq (2 \cdot i_2 + 1)d_G(u,v) \\
&\leq (2 (k-1) + 1)d_G(u,v) \\
&\leq (2k-1) d_G(u,v)
\end{align*}
\label{lem:uvfeasterminates}
\end{lem}

@@

\begin{thm}
For an integer $k\geq 2$, a $(2k - 1)$-approximate distance oracle of G of size $O(kn^{1+1/k})$ and $O(\log k)$ query time can be constructed in $O\left( \min \left\lbrace kmn^{1+\frac{1}{k}}, \sqrt{k}m + kn^{1+c/k}\right\rbrace\right)$ time for some constant $c$.
\end{thm}
{\it Proof.} For variable $k\geq 3$, for each $u$ the bunch $B_u$ in computed in $O(kmn^{2})$ using \cite{Thorup2005}. The remaining preprocessing for $u$ is determining the necessary subsequences of $\mathcal{I}$ and the $j$ that maximized $\delta_j(u)$ for each subsequence.
The necessary subsequences are built

To bound the time to identify these desired $j$, a simple linear search requires $O(k\log k)$ time. This can be improved to $O(k)$ by constructing of a recursive tree $\mathcal{T}_u$ for each $u$ that represents the recursion of $bdist_k$ on the subsequence $\mathcal{I}'$ consisting of all even numbers in $\mathcal{I}$. A bottom-up seach of $\mathcal{T}_u$ can identify for each canonical subsequence (canonical meaning it is a node in $\mathcal{T}$) $i_1,i_1+2,...,i_2$ an index $j$ in the subsequence that maximizes $\delta_j(u)$ in $O(k)$ time. The bottom-up traversal of $T$ requires checking at most $O(\log k)$ subsequences; consider the subsequence $i_1,..,i_2$ in seperate leaf nodes as seen in Fig. \ref{tree}. The path $P$ (red) starting in $p(l_1)$ and ending in $p(l_2)$ has $O(\log k)$ children, which we denote $X$. The union of $X$ clearly covers the entire subsequence $i_1,i_1+2,...,i_2-2,i_2$, and finding a $j$ that maximizes $\delta_j(u)$ in $X$ therefore gives a $j$ that maximizes $\delta_j(u)$ in the subsequence $i_1,...,i_2$. If $j$ is known for each node in $\mathcal{T}$, it follows that finding the desired $j$ for any canonical subsequences is bound by $O(\log k)$\\

The recursion tree for $bdist_k$ for $u$ will have $\theta(k/\log k)$ children and $\theta(k/\log k)$ non-child nodes. Computing $j$ for each subsequence in $T$ will take $O(\log k)$ time per node, so total preprocessing time for $u$ is $O(k)$.

>>>>

Furthermore $d_{G}(p_{i_1}(u),u) \leq i_1 \cdot d_G(u,v)$. When $p_j(u) = w \not\in B_v$
we have $d_G(v,p_{j+1}(v)) \leq d_G(v,p_{j}(u)) \leq d_G(v,u) + d_G(u,p_j(u))$ by Lemma \ref{lem:onedeltaperiteration} and the triangle inequality, as visualized in Fig. \ref{fig:j1}. Hence each iteration of $dist_k(u,v,i)$ increases $d_G(u,w)$ by at most $d_G(u,v)$. At termination we have
\begin{align*}
d_G(u,w) + d_G(w,v) &\leq 2d_G(u,w) + d_G(u,v) \\
&\leq 2(i_1 \cdot d_G(u,v) + (i_2 - i_1)d_G(u,v)) + d_G(u,v) \\
&\leq 2 \cdot i_2 \cdot d_G(u,v) + d_G(u,v) \\
&\leq (2 \cdot i_2 + 1)d_G(u,v) \\
&\leq (2 (k-1) + 1)d_G(u,v) \\
&\leq (2k-1) d_G(u,v)
\end{align*}
\label{lem:uvfeasterminates}
\end{lem}

@@

\begin{thm}
For an integer $k\geq 2$, a $(2k - 1)$-approximate distance oracle of G of size $O(kn^{1+1/k})$ and $O(\log k)$ query time can be constructed in $O\left( \min \left\lbrace kmn^{1+\frac{1}{k}}, \sqrt{k}m + kn^{1+c/k}\right\rbrace\right)$ time for some constant $c$.

\paragraph{Proof.}

For variable $k\geq 3$, for each $u$ the bunch $B_u$ in computed in $O(kmn^{2})$ using \cite{Thorup2005}. The remaining preprocessing for $u$ is determining the necessary subsequences of $\mathcal{I}$ and the $j$ that maximized $\delta_j(u)$ for each subsequence.
For a subsequence $I'=i_1,...i_2$, the necessary subsequences are constructed by finding an even index $i$ that equally splits $I'$ into two sets $i_1,...,i$ and $i,...,i_2$. Then, find a $j$ in $i_1,...,i$ that maximized $\delta_j(u)$. Then recurse on $i_1,...,i$ and $i,...,i_2$. The recursion stop when a subsequence has length at most $\log k$. This recursion corresponds to the recursion of $bdist_k$, and it will henceforth be referred to as the recursion of $bdist_k$.

The query time is bounded by the number of recursions needed before reaching a $(u,v)-feasible$ subsequence of size at most $\log k$. $bdist_k$ uses a binary search, so that at an arbitraty iteration the subsequence will be split in half into subsequences $i_1,...,i$ and $i,...,i_2$. Only the one of these two subsequences that is $(u,v)$-terminal is then recursed upon. As the number of recursions in this binary search is $O(\log k)$ and $i$ and $j$ are known for all subsequences, the query time is $O(\log k)$.
To bound the time to identify these desired $j$, a simple linear search requires $O(k\log k)$ time. This can be improved to $O(k)$ by constructing of a recursive tree $\mathcal{T}_u$ for each $u$ that represents the recursion of $bdist_k$ on the subsequence $\mathcal{I}'$ consisting of all even numbers in $\mathcal{I}$. A bottom-up seach of $\mathcal{T}_u$ can identify for each canonical subsequence (canonical meaning it is a node in $\mathcal{T}$) $i_1,i_1+2,...,i_2$ an index $j$ in the subsequence that maximizes $\delta_j(u)$ in $O(k)$ time. The bottom-up traversal of $T$ requires checking at most $O(\log k)$ subsequences; consider the subsequence $i_1,..,i_2$ in seperate leaf nodes as seen in Fig. \ref{tree}. The path $P$ (red) starting in $p(l_1)$ and ending in $p(l_2)$ has $O(\log k)$ children, which we denote $X$. The union of $X$ clearly covers the entire subsequence $i_1,i_1+2,...,i_2-2,i_2$, and finding a $j$ that maximizes $\delta_j(u)$ in $X$ therefore gives a $j$ that maximizes $\delta_j(u)$ in the subsequence $i_1,...,i_2$. If $j$ is known for each node in $\mathcal{T}$, it follows that finding the desired $j$ for any canonical subsequences is bound by $O(\log k)$\\

The recursion tree for $bdist_k$ for $u$ will have $\theta(k/\log k)$ children and $\theta(k/\log k)$ non-child nodes. Computing $j$ for each subsequence in $T$ will take $O(\log k)$ time per node, so total preprocessing time for $u$ is $O(k)$. Thus, the preprocessing time for finding $j$ for $G$ is $O(nk)\leq O(km)\leq O(kmn^{1/k})$ as desired.
\end{thm}
